---
title: "Which rivers to include?"
author: "Jóhannes Guðbrandsson"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
source("veidi.R")
library(leaflet)
```
# Disclamers
* We don't distinguish between 0 salmon catch and no report. This will inflate the mean for rivers with sporadic salmon catches. 
* We are not using information on river closed to fishing (protected rivers)

# Rivers close to 40-50 cutoff
## In 2020 assessment. 10 year mean higher than 40 and lower than 50

```{r}
ahaetta |>
  filter(expMedal10< 50 & expMedal10 > 40, ath!="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[ahaetta$expMedal10< 50 & ahaetta$expMedal10 > 40 & ahaetta$ath!="ny"]) |>
  mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```

```{r out.width="100%"}
ahaetta |>
  filter(expMedal10< 50 & expMedal10 > 40, ath!="ny") |>
  mutate(merki = paste(adal_nr, nafn)) |>
  leaflet() |>
  addTiles() |>
  addLabelOnlyMarkers(lng = ~V, lat = ~N, label = ~merki,
                      labelOptions = labelOptions(noHide = T, textsize = "12px"),
                      clusterOptions = markerClusterOptions())

```

## In 2020 assessment. 10 year mean lower than 40, longterm mean higher than 40
* Ósá / Setbergsá catch reports not returned in latest years. Records for Stóru Langadalsá returned but not for Setbergsá
* Selá temperature dependent? 
* Selfljót / Gilsá. Temperature dependent?
* Vatnsholtsá / Vatnasvæði Lýsu. Mostly trout / charr fishing. Catch and release in latest years

```{r}
ahaetta |>
  filter(expMedal10< 40 & expMedal > 40, ath!="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[ahaetta$expMedal10< 40 & ahaetta$expMedal > 40 & ahaetta$ath!="ny"]) |>
  mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```

```{r out.width="100%"}
ahaetta |>
  filter(expMedal10< 40 & expMedal > 40, ath!="ny") |>
  mutate(merki = paste(adal_nr, nafn)) |>
  leaflet() |>
  addTiles() |>
  addLabelOnlyMarkers(lng = ~V, lat = ~N, label = ~merki,
                      labelOptions = labelOptions(noHide = T, textsize = "12px"),
                      clusterOptions = markerClusterOptions())

```

## Not in 2020 assessment. 10 year mean lower than 40, longterm mean higher than 40
* Effect of smolt releases
* Most likely a wild stock in Selá

```{r}
ahaetta |>
  filter(expMedal10< 40 & expMedal > 40 & ath=="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[ahaetta$expMedal10< 40 & ahaetta$expMedal > 40 & ahaetta$ath=="ny"]) |>
  mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  xlab("") +
  ylab("Estimated stock size") +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```

```{r out.width="100%"}
ahaetta |>
  filter(expMedal10< 40 & expMedal > 40 & ath=="ny") |>
  mutate(merki = paste(adal_nr, nafn)) |>
  leaflet() |>
  addTiles() |>
  addLabelOnlyMarkers(lng = ~V, lat = ~N, label = ~merki,
                      labelOptions = labelOptions(noHide = T, textsize = "12px"),
                      clusterOptions = markerClusterOptions())

```

## Not in 2020 assessment. 10 year mean not available, longterm mean higher than 40
* Most of these smolt releases / adult salmon transport
* Móra, Tjarnará, Kiðafellsá, Hoffelssá (Laxá í Nesjum) with wild stock? Stock size?

```{r}
ahaetta |>
  filter(is.na(expMedal10) & expMedal > 40 & ath=="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[is.na(ahaetta$expMedal10) & ahaetta$expMedal > 40 & ahaetta$ath=="ny"]) |> 
  #mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```

* Need coordinates

#  Not in 2020 assessment. 10 year over 50

* Smolt releases / ranching. Shoud be in the model, but will not affect production if they go over 4% intrusion
  * Affal, Skógá, Lagarfljót, (Jökla?)
* New stocks. Fish ladders and hydropower
  * Svínafossá, (Jökla?)
* Kúðafljót as Tungufljót in 2020 assessment
* Skaftá as Geirlandsá in 2020 assessment
* Don't know why the rest were not in the 2020 assessment
  * Staðará Súgandafirði only 7 years of records, Lagarfljót 9 years
  
```{r}
ahaetta |>
  filter(expMedal10 >= 50, ath=="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[ahaetta$expMedal10 >= 50 & ahaetta$ath=="ny"]) |>
  mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```

```{r out.width="100%"}
ahaetta |>
  filter(expMedal10 >= 50, ath=="ny") |>
  mutate(merki = paste(adal_nr, nafn)) |>
  leaflet() |>
  addTiles() |>
  addLabelOnlyMarkers(lng = ~V, lat = ~N, label = ~merki,
                      labelOptions = labelOptions(noHide = T, textsize = "12px"),
                      clusterOptions = markerClusterOptions())

```

# In 2020 assessment. 10 year mean lower than 40, longterm mean lower than 40

```{r}
ahaetta |>
  filter(expMedal10< 40 & expMedal < 40 & ath!="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[ahaetta$expMedal10< 40 & ahaetta$expMedal < 40 & ahaetta$ath!="ny"]) |>
  mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```


```{r out.width="100%"}
ahaetta |>
  filter(expMedal10< 40 & expMedal < 40 & ath!="ny") |>
  mutate(merki = paste(adal_nr, nafn)) |>
  leaflet() |>
  addTiles() |>
  addLabelOnlyMarkers(lng = ~V, lat = ~N, label = ~merki,
                      labelOptions = labelOptions(noHide = T, textsize = "12px"),
                      clusterOptions = markerClusterOptions())

```

# Not in 2020 assessment. 10 year mean lower than 40, longterm mean lower than 40
* Some database errors in the 80's need to be checked

```{r}
ahaetta |>
  filter(expMedal10<40 & expMedal < 40 & ath=="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[ahaetta$expMedal10<40 & ahaetta$expMedal < 40 & ahaetta$ath=="ny"]) |> 
  mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```

# Not in 2020 assessment. 10 year mean not available, longterm mean lower than 40
```{r}
ahaetta |>
  filter(is.na(expMedal10) & expMedal < 40 & ath=="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[is.na(ahaetta$expMedal10) & ahaetta$expMedal < 40 & ahaetta$ath=="ny"]) |> 
  #mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```


# In 2020 assessment. 10 year mean over 50
```{r}
ahaetta |>
  filter(expMedal10>50 & ath!="ny") |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r out.width="100%"}
stofnmat |>
  filter(adal_nr %in% ahaetta$adal_nr[ahaetta$expMedal10>50 & ahaetta$ath!="ny"]) |> 
  mutate(expMedal10 = ifelse(ar<2014,NA,expMedal10)) |>
  #filter(expMedal10<100) |>
  pivot_longer(starts_with("stofn_tmp"),names_to = "sleppt_kodi", values_to = "stofn_tmp", names_prefix = "stofn_tmp_") |> 
  #filter(!(adal_nr %in% c("Bjarnafjarðará"=152))) |>
  ggplot(aes(ar,stofn_tmp, fill=factor(sleppt_kodi))) +
  geom_bar(stat = "identity")+
  geom_path(aes(y=expMedal10), col="black") +
  geom_path(aes(y=expMedal), col="black", linetype = 2) +
  facet_wrap(~paste(adal_nr,adalvatnsfall), scales = "free_y") +
  xlab("") +
  ylab("Estimated stock size") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position = "none")
```


```{r out.width="100%"}
ahaetta |>
  filter(expMedal10>50 & ath!="ny") |>
  mutate(merki = paste(adal_nr, nafn)) |>
  leaflet() |>
  addTiles() |>
  addLabelOnlyMarkers(lng = ~V, lat = ~N, label = ~merki,
                      labelOptions = labelOptions(noHide = T, textsize = "12px"),
                      clusterOptions = markerClusterOptions())

```