---
title: "Population size variation"
author: "Jóhannes Guðbrandsson"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", warning = FALSE, message = FALSE)
```

# Goal / questions

1.  What probability distribution should we use?
2.  How should we estimate parameters? Is the variation different between rivers?
3.  Do we need to take temporal correlation into account?

# Counters
## Data
```{r}
library(tidyverse)
library(modelr)
library(zoo)
library(patchwork)
library(mar)

mar <- connect_mar()
default_theme <- theme_set(theme_bw())

gogn <- read_delim("teljaragogn.csv") |>
  filter(!is.na(fjoldi)) |>
  filter(!vatnsfall %in% c("Selá","Laugardalsá")) |>
  #filter(!(ar < 1992 & vatnsfall=="Blanda")) |>
  mutate(ln = log(fjoldi)) |>
  group_by(vatnsfall) |>
  mutate( m5 = lag(rollmean(ln, 5,fill=NA, align = "right"))) |>
  ungroup()
```

```{r out.width="100%"}
gogn |>
  ggplot(aes(ar, fjoldi, col=vatnsfall)) +
  geom_point() +
  geom_path() +
  facet_wrap(~vatnsfall, scales = "free") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) 
```

```{r out.width="100%"}
gogn |>
  ggplot(aes(ar, fjoldi, col=vatnsfall)) +
  geom_point() +
  geom_path() +
  facet_wrap(~vatnsfall, scales = "free") +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_y_log10()
```

```{r}
gogn |>
  group_by(vatnsfall) |>
  summarise(start=min(ar), stop=max(ar), fjoldi_ara=n(), medal_fjoldi=mean(fjoldi), log_m_trans=exp(mean(ln)), log_medal= mean(ln), stadalfravik= sd(fjoldi), log_stadal=sd(ln)) |>
  arrange(medal_fjoldi) |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r}
gogn |>
  group_by(vatnsfall) |>
  summarise(start=min(ar), stop=max(ar), fjoldi_ara=n(), medal_fjoldi=mean(fjoldi), log_m_trans=exp(mean(ln)), log_medal= mean(ln), stadalfravik= sd(fjoldi), log_stadal=sd(ln)) |>
  ggplot(aes(log_medal, log_stadal))+
  geom_point() +
  geom_smooth(method = "lm")
```

# Fish counters - distribution
```{r}
mNorm <- lm(fjoldi~vatnsfall, data=gogn) #Meðaltöl í fjölda
mLNorm <- lm(ln~vatnsfall, data=gogn) #Meðaltöl á log
mLag <- lm(ln~vatnsfall + m5, data=gogn)
mLag2 <- lm(ln~vatnsfall + vatnsfall*m5, data=gogn)

gogn |>
  spread_residuals(mNorm,mLNorm, mLag, mLag2) ->
  gogn
```
## Normal distribution
```{r}
plot(mNorm)
```

```{r}
shapiro.test(gogn$mNorm)
```

## Log-normal
```{r}
plot(mLNorm)
```
```{r}
shapiro.test(gogn$mLNorm)
```

# Fisktalning - Minni
```{r}
mLNorm <- lm(ln~vatnsfall, data=filter(gogn,!is.na(m5)))
anova(mLNorm, mLag)
anova(mLag, mLag2)
```

## Partial autocorrelation of residuals from ANOVA
```{r}
for (v in unique(gogn$vatnsfall)){
  #acf(gogn$mLNorm[gogn$vatnsfall==v], main=v)
  pacf(gogn$mLNorm[gogn$vatnsfall==v], main=v)
}
```

## Partial autocorrelation of residuals from ANCOVA (5-year running mean)
```{r}
for (v in unique(gogn$vatnsfall)){
  #acf(na.omit(gogn$mLag[gogn$vatnsfall==v]), main=v)
  pacf(na.omit(gogn$mLag[gogn$vatnsfall==v]), main=v)
}
```

## Standard deviation
```{r}
summary(mLNorm)$sigma
summary(mLag)$sigma
```

# Salmon catches
## Data
```{r}
listi <- c("Andakílsá"=12,"Blanda"=158,"Elliðaár"=411,"Norðurá"=1861, "Krossá"=1462,
           "Langá"=1568,"Miðfjarðará"=1733, "Hítará"=956,"Sog"=2271,"Stóra-Laxá"=2339,
           "Ásgarðslækur"=284,"Skaftá"=2164,"Héraðsvötn"=955,
           "LaxáLeir" = 1589,"LaxáAðal"=1591,"Þverá"=2870,"LaxáKjós"=1590, "Álftá"=69,
           "Fnjóská"=537,"Grímsá"=749,"Haffjarðará"=835,
           "Hafralónsá"=851,"Haukadalsa"=883,"Hofsa"=996,"Laugardalsa"=1579,
           "Svalbarðsa"=2424, "Brynjudalsá"=257,"FáskrúðDölum"=465,"Hrútafjarðará" =1106,
           "HölknáÞist"=1216, "LaxáDölum"=1600,"Leirvogsá"=1615, "SeláVopn"=2080,
           "StaðaráSt"=2282)
veidibok_veidibok(mar) |> 
  left_join(tbl_mar(mar,"ops$johannes.VEIDIBOK_TO_WATER"), by=c("vatnsfall_id"="id")) |>
  left_join({fv_vatn(mar) |>
      select(vatn_nr, adalvatnsfall_nr, vatnsfall_til_nr)}, by=c("water_no"="vatn_nr")) |>
  left_join({fv_vatn(mar) |> 
      select(vatn_nafn, vatn_nr) |> 
      rename(adalvatnsfall = vatn_nafn)}, by= c("adalvatnsfall_nr"="vatn_nr")) |>
  filter(fisktegund_id == 91, ar <2023, ar > 1973) |>
  filter(adalvatnsfall_nr %in% listi |
           water_no %in% listi) |>
  filter(!(water_no==1568 & ar %in% 1989:1990),
         !(adalvatnsfall_nr==996 & ar==2008),
         vatnsfall!="Blöndulón",
         !(water_no==2574 & ar %in% 2008:2009), #Tungulækur hafbeit
         water_no!=922) |> #Hellisá flutningur á fiski
  #replace_na(list(sleppt_kodi=0)) |>
  group_by(adalvatnsfall_nr) |>
  collect(n=Inf) |>
  mutate(vatnsfall= str_flatten(sort(unique(vatnsfall)), collapse = " ")) |> 
  ungroup() |>
  count(vatnsfall, adalvatnsfall_nr, ar) |> 
  mutate(ln=log(n)) |>
  arrange(adalvatnsfall_nr, ar) |>
  group_by(adalvatnsfall_nr) |>
  collect(n=Inf) |>
  mutate( mInf = mean(ln,na.rm = T),
          m5 = lag(rollmean(ln, 5,fill=NA, align = "right")),
          m10 = lag(rollmean(ln, 10,fill=NA, align = "right"))) |>
  ungroup() ->
  test
```

```{r}
test |>
  ggplot(aes(ar,n, col=vatnsfall)) +
  geom_path() +
  #geom_point() +
  #geom_path(aes(y=m5), linetype = 2) +
  #geom_path(aes(y=m10), linetype = 3) +
  facet_wrap(~adalvatnsfall_nr, scales="free_y") +
  theme(legend.position = "none")
```

```{r}
test |>
  ggplot(aes(ar,n, col=vatnsfall)) +
  geom_path() +
  #geom_point() +
  #geom_path(aes(y=m5), linetype = 2) +
  #geom_path(aes(y=m10), linetype = 3) +
  facet_wrap(~adalvatnsfall_nr, scales="free_y") +
  theme(legend.position = "none") +
  scale_y_log10()
```

```{r}
test |>
  group_by(adalvatnsfall_nr,vatnsfall) |>
  summarise(start=min(ar), stop=max(ar), fjoldi_ara=n(), medal_fjoldi=mean(n), log_m_trans=exp(mean(ln)), log_medal= mean(ln), stadalfravik= sd(n), log_stadal=sd(ln)) |>
  #arrange(medal_fjoldi) |>
  kableExtra::kbl(digits = 2) |>
  kableExtra::kable_styling("striped", full_width = F) 
```

```{r}
test |>
  group_by(vatnsfall) |>
  summarise(start=min(ar), stop=max(ar), fjoldi_ara=n(), medal_fjoldi=mean(n), log_m_trans=exp(mean(ln)), log_medal= mean(ln), stadalfravik= sd(n), log_stadal=sd(ln)) |>
  ggplot(aes(log_medal, log_stadal))+
  geom_point() +
  geom_smooth(method = "lm",se=FALSE) +
  geom_smooth(col="black", se=FALSE)
```

# Catch data distribution
Skip checking the normal distribution (it's obviously does not fit)

## log-normal
```{r}
#Líkön
mL <- lm(ln~vatnsfall, (test))
mLm5 <- lm(ln~vatnsfall + m5, data = (test))
mLm10 <- lm(ln~vatnsfall + m10, data = test)
mLNA <- lm(ln~vatnsfall, na.omit(test))
mLm5NA <- lm(ln~vatnsfall + m5, data = na.omit(test))
mLm5m10 <- lm(ln~vatnsfall + m5+m10, data = test)
mL0m5 <- lm(ln~m5, data = na.omit(test))
mL0m5m10 <- lm(ln~m5 + m10, data = na.omit(test))
mL0m10 <- lm(ln~m10, data = na.omit(test))
mLm5NAspec <- lm(ln~vatnsfall + vatnsfall*m5, data = na.omit(test))

test |>
  spread_residuals(mL, mLm5,mLm10,mLm5m10) ->
  test
```

```{r}
plot(mL)
```

```{r}
shapiro.test(test$mL)
shapiro.test(test$mLm5m10)
```

Feitir halar, veiðihlutfall og léleg skráning?

# Catch data lag 5 year, 10 year
```{r}
drop1(mLm5m10, test="F")
drop1(mLm5NAspec, test = "F")
```

5 year mean with highest F-value

## Standard deviation
```{r}
summary(mLm5m10)$sigma
summary(mLm5)$sigma
summary(mL)$sigma
```

## Partial autocorrelation 5 year lag
```{r}
for (v in unique(test$vatnsfall)){
  #acf(na.omit(test$m2[test$vatnsfall==v]), main=v)
  pacf(na.omit(test$mLm5[test$vatnsfall==v]), main=v)
}
```

# Results
1.  What probability distribution should we use?
    - log-Normal distribution seems to adequate. Might be worth it to take a closer look at smaller stocks
2.  How should we estimate parameters? Is the variation different between rivers?
    - We should work on log-scale, geometric mean instead of arithmetic. The standard deviation seems to similar among river (0.5). More variation for small stocks biological or a "measurement error"?
3.  Do we need to take temporal correlation into account?
    - Strong correlation with 5 year mean and 10 year mean also significant. One year auto-correlation in some rivers (2SW rivers)
    - 10 year mean used in the previous edition of the risk assessment
